### 2011-02-03 docs/architecture/architecture.tex
more making things pretty

A basic {\tt GNUmakefile} begins with:
\begin{lstlisting}[language=bash,mathescape=false]
  NDEBUG := t
  MPI    :=
  OMP    :=
\end{lstlisting}
Here, {\tt NDEBUG} is true if we are building an optimized executable.
Otherwise, the debug version is built---this typically uses less
optimization and adds various runtime checks through compiler flags.
{\tt MPI} and {\tt OMP} are set to true if we want to use either MPI
or OpenMP for parallelization.  If {\tt MPI := t}, you will need to
have the MPI libraries installed, and their location may need to be 
specified in {\tt MAESTRO/mk/GMakeMPI.mak}.

The next line sets the compiler to be used for compilation:
\begin{lstlisting}[language=bash,mathescape=false]
  COMP := Intel
\end{lstlisting}
The MAESTRO build system knows what options to use for various
compiler families.  The {\tt COMP} flag specifies which compiler to
use.  Allowed values include {\tt Intel}, {\tt gfortran}, {\tt PGI},
and {\tt PathScale}.  The specific details of these choices are
defined in the {\tt MAESTRO/mk/comps/} directory.

{\tt MKVERBOSE} set to true will echo the build commands to the
terminal as the are executed.
\begin{lstlisting}[language=bash,mathescape=false]
  MKVERBOSE := t
\end{lstlisting}

The next two lines define where the top of the source tree is
include the basic Makefile information (like compiler
flags).
\begin{lstlisting}[language=bash,mathescape=false]
  FPARALLEL := ../..

@@

A MAESTRO application is built from several pacakges (the
multigrid solver, an EOS, a reaction network, etc.).  The core
MAESTRO packages are always included, so a problem only needs
to define the EOS, reaction network, and conductivities to
use, as well as any extra, problem-specific files.  
\begin{lstlisting}[language=bash,mathescape=false]
EOS_DIR := extern/EOS/helmeos   
CONDUCTIVITY_DIR := extern/conductivity/constant
NETWORK_DIR := extern/networks/ignition

@@

Runtime parameters listed in the {\tt MAESTRO/\_parameters} file are
parsed at compile time and the file {\tt probin.f90} is written and
compiled.  This is a Fortran module that holds the values of the
runtime parameters and makes them available to any routine.  The
variable {\tt PROBIN\_PARAMETERS} should list all of the files that
contain runtime parameters for the current problem.  You must always
include the main {\tt ../\_parameters} file (i.e.\ {\tt
  MAESTRO/\_parameters}).  Sometimes a problem defines their own
runtime parameters (for instance {\tt \_parameters.myproblem}), in which
case, its parameter file should also appear in {\tt
  PROBIN\_PARAMETERS} (see \S~\ref{sec:def_runtime_param}).
\begin{lstlisting}[language=bash,mathescape=false]
  PROBIN_PARAMETERS := ../_parameters
\end{lstlisting}

The final line in the {\tt GNUmakefile} includes the rules to actually
build the executable.
\begin{lstlisting}[language=bash,mathescape=false]
  include $(FPARALLEL)/MAESTRO/GMaestro.mak
\end{lstlisting}
%$

@@

\begin{verbatim}
  subroutine example(s,dx,dt)

    use bl_types
    use multifab_module
    use geometry, only: dm, nlevs
    use variables, only: rho_comp
\end{verbatim}

@@

\begin{verbatim}
    type(multifab) , intent(inout) :: s(:)
    real(kind=dp_t), intent(in   ) :: dx(:,:),dt
\end{verbatim}

@@

\begin{verbatim}
    ! Local variables
    real(kind=dp_t), pointer :: sp(:,:,:,:)
    integer :: i,n,ng_sp
    integer :: lo(dm),hi(dm)
\end{verbatim}

@@

\begin{verbatim}
    ng_sp = nghost(s(1))
\end{verbatim}

@@

\begin{verbatim}
    do n=1,nlevs
       do i = 1, s(n)%nboxes
\end{verbatim}

@@

\begin{verbatim}
          if ( multifab_remote(s(n), i) ) cycle
          sp => dataptr(s(n), i)
          lo =  lwb(get_box(s(n), i))
          hi =  upb(get_box(s(n), i))
\end{verbatim}

@@

\begin{verbatim}
          select case (dm)
          case (2)
             call example_2d(sp(:,:,1,rho_comp),ng_sp,lo,hi,dx(n,:),dt)
          case (3)
             call example_3d(sp(:,:,:,rho_comp),ng_sp,lo,hi,dx(n,:),dt)
          end select
       enddo    ! end loop over boxes

@@

  end subroutine example
\end{verbatim}

@@

\begin{verbatim}
  subroutine example_2d(density,ng_sp,lo,hi,dx,dt)

@@

  end subroutine example_2d
\end{verbatim}

@@

The three-dimensional case is similar, with the {\tt density} array
declared as 
\begin{verbatim}
  density(lo(1)-ng_sp:,lo(2)-ng_sp:,lo(3)-ng_sp)
\end{verbatim}
and an additional loop over the `z' coordinate (from {\tt lo(3)} to
{\tt hi(3)}).

>>>>

A basic {\tt GNUmakefile} begins with:
\begin{lstlisting}[language={[gnu]make},mathescape=false]
  NDEBUG := t
  MPI    :=
  OMP    :=
\end{lstlisting}
Here, {\tt NDEBUG} is true if we are building an optimized executable.
Otherwise, the debug version is built---this typically uses less
optimization and adds various runtime checks through compiler flags.
{\tt MPI} and {\tt OMP} are set to true if we want to use either MPI
or OpenMP for parallelization.  If {\tt MPI := t}, you will need to
have the MPI libraries installed, and their location may need to be 
specified in {\tt MAESTRO/mk/GMakeMPI.mak}.

The next line sets the compiler to be used for compilation:
\begin{lstlisting}[language={[gnu]make},mathescape=false]
  COMP := Intel
\end{lstlisting}
The MAESTRO build system knows what options to use for various
compiler families.  The {\tt COMP} flag specifies which compiler to
use.  Allowed values include {\tt Intel}, {\tt gfortran}, {\tt PGI},
and {\tt PathScale}.  The specific details of these choices are
defined in the {\tt MAESTRO/mk/comps/} directory.

{\tt MKVERBOSE} set to true will echo the build commands to the
terminal as the are executed.
\begin{lstlisting}[language={[gnu]make},mathescape=false]
  MKVERBOSE := t
\end{lstlisting}

The next two lines define where the top of the source tree is
include the basic Makefile information (like compiler
flags).
\begin{lstlisting}[language={[gnu]make},mathescape=false]
  FPARALLEL := ../..

@@

A MAESTRO application is built from several pacakges (the
multigrid solver, an EOS, a reaction network, etc.).  The core
MAESTRO packages are always included, so a problem only needs
to define the EOS, reaction network, and conductivities to
use, as well as any extra, problem-specific files.  
\begin{lstlisting}[language={[gnu]make},mathescape=false]
EOS_DIR := extern/EOS/helmeos   
CONDUCTIVITY_DIR := extern/conductivity/constant
NETWORK_DIR := extern/networks/ignition

@@

Runtime parameters listed in the {\tt MAESTRO/\_parameters} file are
parsed at compile time and the file {\tt probin.f90} is written and
compiled.  This is a Fortran module that holds the values of the
runtime parameters and makes them available to any routine.  The
variable {\tt PROBIN\_PARAMETERS} should list all of the files that
contain runtime parameters for the current problem.  You must always
include the main {\tt ../\_parameters} file (i.e.\ {\tt
  MAESTRO/\_parameters}).  Sometimes a problem defines their own
runtime parameters (for instance {\tt \_parameters.myproblem}), in which
case, its parameter file should also appear in {\tt
  PROBIN\_PARAMETERS} (see \S~\ref{sec:def_runtime_param}).
\begin{lstlisting}[language={[gnu]make},mathescape=false]
  PROBIN_PARAMETERS := ../_parameters
\end{lstlisting}

The final line in the {\tt GNUmakefile} includes the rules to actually
build the executable.
\begin{lstlisting}[language={[gnu]make},mathescape=false]
  include $(FPARALLEL)/MAESTRO/GMaestro.mak
\end{lstlisting}
%$

@@

\begin{lstlisting}[language={[95]fortran},mathescape=false]
  subroutine example(s,dx,dt)

    use bl_types
    use multifab_module
    use geometry, only: dm, nlevs
    use variables, only: rho_comp
\end{lstlisting}

@@

\begin{lstlisting}[language={[95]fortran},mathescape=false]
    type(multifab) , intent(inout) :: s(:)
    real(kind=dp_t), intent(in   ) :: dx(:,:),dt
\end{lstlisting}

@@

\begin{lstlisting}[language={[95]fortran},mathescape=false]
    ! Local variables
    real(kind=dp_t), pointer :: sp(:,:,:,:)
    integer :: i,n,ng_sp
    integer :: lo(dm),hi(dm)
\end{lstlisting}

@@

\begin{lstlisting}[language={[95]fortran},mathescape=false]
    ng_sp = nghost(s(1))
\end{lstlisting}

@@

\begin{lstlisting}[language={[95]fortran},mathescape=false]
    do n=1,nlevs
       do i = 1, s(n)%nboxes
\end{lstlisting}

@@

\begin{lstlisting}[language={[95]fortran},mathescape=false]
          if ( multifab_remote(s(n), i) ) cycle
          sp => dataptr(s(n), i)
          lo =  lwb(get_box(s(n), i))
          hi =  upb(get_box(s(n), i))
\end{lstlisting}

@@

\begin{lstlisting}[language={[95]fortran},mathescape=false]
          select case (dm)
          case (2)
             call example_2d(sp(:,:,1,rho_comp),ng_sp,lo,hi,dx(n,:),dt)
          case (3)
             call example_3d(sp(:,:,:,rho_comp),ng_sp,lo,hi,dx(n,:),dt)
          end select
       enddo    ! end loop over boxes

@@

  end subroutine example
\end{lstlisting}

@@

\begin{lstlisting}[language={[95]fortran},mathescape=false]
  subroutine example_2d(density,ng_sp,lo,hi,dx,dt)

@@

  end subroutine example_2d
\end{lstlisting}

@@

The three-dimensional case is similar, with the {\tt density} array
declared as 
\begin{lstlisting}[language={[95]fortran},mathescape=false]
  density(lo(1)-ng_sp:,lo(2)-ng_sp:,lo(3)-ng_sp:)
\end{lstlisting}
and an additional loop over the `z' coordinate (from {\tt lo(3)} to
{\tt hi(3)}).
